Efficient and Robust Plane-Based Boolean Operations on N Meshes

### 2016-12-12 paper.tex
here!!!!

\usepackage{enumitem}
\usepackage{algorithm}
\usepackage{multirow}
\usepackage{slashbox}
%\usepackage{ulem}
%\usepackage[numbers,sort&compress]{natbib}
\renewcommand{\arraystretch}{1.3}

\usepackage{color}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\newcommand{\tabincell}[2]{\begin{tabular}{@{}#1@{}}#2\end{tabular}}
\floatname{algorithm}{Procedure}

\begin{document}
\title{Efficient Non-incremental Constructive Solid Geometry Evaluation for Triangular Meshes}
\author{Rui~Wang,~
        Xudong~Jiang,~
        Hongbo~Fu,~
        Bin~Sheng,~and
        Enhua~Wu
\IEEEcompsocitemizethanks{

@@

\IEEEtitleabstractindextext{
\begin{abstract}
\end{abstract}

\begin{IEEEkeywords}
Boolean operations, triangle mesh, CSG evaluation, plane-based geometry.
\end{IEEEkeywords}

@@

\IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
\IEEEPARstart{C}{onstructive} solid geometry (CSG) has long been a popular modeling tool of computer-aided design and computer-aided manufacturing (CAD/CAM). It constructs complex models by combining primitives using a series of regularized Boolean operations \cite{requicha1977mathematical}: union, intersection and difference. CSG is often represented by a binary tree, called a CSG tree, whose leaves are primitives and whose internal nodes are Boolean operations.
\newline

{\color{red}{
Main Contribution:

1. We develop a united representation of intersection lines between CSG primitives. Under this representation, Face re-tessellation and classification can be done robustly in any degenerate cases by our pipeline, while keeping good performance.

2. Unlike any of the previous method, during the algorithm, we conduct point-in-polyhedron tests only on single seed vertex. We develop a propagating to spread the classification information from the seed vertex into the whole CSG. In addition, because of this, we do not need to build a full space division to accelerate point-in-polyhedron tests, which is usually time-consuming.
}}

@@

\begin{figure*}[!t]
\centering
\includegraphics[width=7.1in]{flowchart}
\caption{An overview of the proposed method for CSG evaluation. In this toy example, the CSG tree represents a solid constructed by (Cube $\cup$ Sphere - Cylinder). (d) shows how to classify faces step by step using a flood filling algorithm. Faces of the same primitive are grouped as the inter-primitive group. Then faces of the same inter-primitive group are further divided into intra-primitive groups according to their conditions of intersection (c). The unvisited faces are black. Different colors of other faces indicate that they are accepted and belong to different intra-primitive groups. Rejected faces are rendered with wire-frame. The seed face of flood filling in each primitive is highlighted in red.}
\label{fig_overview}
\end{figure*}

CSG evaluation has had notorious problems with robustness since its inception in the 1980s \cite{requicha1985boolean, laidlaw1986constructive}. The non-robustness is inherited from the building blocks of CSG: Boolean operations on solids. Many researchers have attempted to solve such an issue using arbitrary precision arithmetic \cite{banerjee1996topologically, fortune1995polyhedral, keyser2004esolid, granados2003boolean, hachenberger2005boolean} and exact interval computation \cite{fang1993robustness, hu1996robust, segal1990using}. However, these methods are often too expensive in computation time and memory to be practical for evaluation of CSG with massive faces. For example, in the Computational Geometry Algorithms Library (CGAL) \cite{cgal:hk-bonp3-15a}, the state-of-the-art robust Boolean operation algorithm \cite{hachenberger2005boolean} (implemented with arbitrary precision arithmetic) is more than 20 times slower than its non-robust version.

Sugihara and Iri \cite{sugihara1990solid} introduced a plane-based representation of polyhedra. They found that Boolean operations are fundamentally robust using the plane equation as the primary geometry representation. In this condition, the evaluation of Boolean operations can be performed with no `constructions' \cite{shewchuk1999lecture}, thereby avoiding the introduction of any numerical errors.

Berstein and Fussell \cite{bernstein2009fast} noted that a binary space partitioning (BSP) merging algorithm in Boolean operations \cite{thibault1987set, naylor1990merging} is actually a plane-based technique. Therefore, they combined the two conceptions---plane-based geometry and BSP merging---to develop an unconditionally robust method for Boolean operations of polyhedra. Using Shewchuk's adaptive geometry predicate technique \cite{shewchuk1997adaptive}, this method can be only twice slower than non-robust methods. Our approach adopts this technique in the stage of face classification to ensure the robustness.

Berstein and Fussell's method does not alleviate the problem of high memory consumption of BSP-based algorithms. Thus, it remains unprepared for polyhedra with massive faces. However, it inspired Campen and Kobbelt to develop a more delicate approach  \cite{campen2010exact}. This approach solves the problem of exact and efficient conversion of polyhedra between vertex-based and plane-based representations. Moreover, by leveraging the adaptive octree, BSP structures are nested into octree cells where the intersection between primitives occurs. For efficiency, face classification is performed using cells as the basic classification units. This localized scheme saves considerable computing resources and retains the topology of non-intersected areas.

Such localized scheme is widely used in different methods of CSG boundary evaluation \cite{pavic2010hybrid, wang2011approximate, feito2013fast, campen2010exact}. These methods are usually based on intersection computation and face membership classification. The success of such localized scheme is enabled by two facts. First, the intersections between primitives are locally distributed. Second, space labels of a face are often coherent with those of its neighborhood. The former makes it possible to reduce the space that requires an intersection test. The latter enables sharing of space labels between adjacent faces to accelerate face classification.

However, most of these methods use the cube (e.g., voxels, octree cells) as the unit of the face group, which is often the natural result of reusing space-division data structures constructed for intersection computation. In this way, connected faces that share the same labels might be distributed in different cubes. Thus, repetitive classifications may be required for each cube. Furthermore, faces of different classifications may coexist within a small space, forming a ‘mixed’ area (e.g., the red areas in Figs. \ref{fig:cluster}(a) and (b)). Because the size of the cubes cannot be illimitably small, special cubes that contain faces with different space labels typically must be constructed in these mixed areas. Processing these special cubes is often complex and time-consuming. To more efficiently share space labels, we therefore use faces, rather than cubes, as the basic grouping units. Grouping is efficiently performed by utilizing geometry connectivity with a flood-filling algorithm.

Recently, Feito et al. \cite{feito2013fast} proposed a method for Boolean evaluation. This method also uses a face-based grouping scheme. To share classification results, faces that do not intersect with other primitives are grouped together according to geometry connectivity. This method and the proposed method have similar stages for single Boolean operation evaluation. One of the major differences is that our method supports non-incremental evaluation of CSG. For this purpose, we designed a series of dedicated stages. Additionally, our face grouping scheme is intended for both non-intersected faces and faces that intersect with other primitives. It thus provides a uniform treatment for all faces and further accelerates the face classification stage. Moreover, the similarity between adjacent groups is omitted by Feito et al.'s method. Although different groups have different space labels, we observe that they may still share some of these labels if they are neighboring. We utilize this kind of space label coherence between groups for better performance.

On the other hand, with the development of general-purpose computing on graphics processing units (GPGPUs), many researchers  \cite{wang2011approximate, zhao2011parallel, museth2002level, chen1999volumetic, eisemann2008single} have tried to utilize the grand computation power of graphics hardware for Boolean operations. These methods often have good performance and are suitable for interactive applications, such as digital sculpting. However, owing to the paralleled features of graphics hardware, these methods are usually voxel-based and support only approximate evaluation that inevitably suffers from geometric detail loss, especially at the intersection areas of primitives.

\section{Overview}

\label{sec:overview}

As in most previous works \cite{feito2013fast, bernstein2009fast}, we require the input primitives of the CSG tree to be watertight Nef polyhedra with manifold surfaces. We assume there is no self-intersection. Because we adopt a flood-filling strategy, we additionally assume that the connectivity of adjacent faces is available. Our current implementation is dedicated to primitives of triangular meshes for simplicity. Our approach efficiently and robustly computes the boundary of CSG solids in three steps, as illustrated in Fig. \ref{fig_overview}. Our main contributions exist in Steps 2 and 3.

\subsection{Adaptive Octree Construction}

As a typical routine of localization, our method uses the adaptive octree to frequently accelerate the used spatial query afterwards. Our octree construction implementation is akin to the implementation in Feito et al.'s method \cite{feito2013fast}. The major difference is that, because our method is for non-incremental CSG evaluation, the octree is constructed on the bounding box that contains all input primitives (Fig. \ref{fig_overview}(b)). To avoid presentation ambiguity between a CSG leaf and an octree leaf, the latter is called a cell. Intersection detection between triangle faces and cells can be efficiently performed by using the separating axis theorem \cite{gottschalk1996obbtree}. Cells are classified into two types: if all triangles that intersect a cell belong to the same primitive, we call it a \emph{non-critical} cell. Otherwise, it is a \emph{critical} cell. This classification is needed for intersection computation in the next step.

\subsection{United Representation of Intersection Construction}
\label{sec:overview:isect}

\subsection{Multilevel Flood-filling Classification}

\section{United Representation of Primitive Intersection}
{\color{red}{
Primitive intersects primitive, generating intersection line segments (Fig. \ref{fig_raw1}).

\begin{figure}
  \centering
  % Requires \usepackage{graphicx}
  \includegraphics[width=3.5in]{raw1}\\
  \caption{triangle contains intersection lines, two-primitive case}\label{fig_raw1}
\end{figure}

Considering degenerate cases in n-primitives CSG can be very complex (Fig. \ref{fig_raw2}).

\begin{figure}
  \centering
  % Requires \usepackage{graphicx}
  \includegraphics[width=3.5in]{raw2}\\
  \caption{triangle contains intersection lines, n-primitives degenerate cases}\label{fig_raw2}
\end{figure}

We converted it into a united representation, making it clear and easy to process by the following process (Fig. \ref{fig_raw3}).

\begin{figure}
  \centering
  % Requires \usepackage{graphicx}
  \includegraphics[width=3.5in]{raw3}\\
  \caption{triangle contains intersection lines, n-primitives degenerate cases, using united representation}\label{fig_raw3}
\end{figure}
}}

\subsection{Plane-based Triangle Intersection}

{\color{red}{why using triangle intersection, rather than edge-face based, etc. --- for coplanar test.}}
\newline\newline

\begin{figure}[t]
\centering
\includegraphics[width=3.0in]{projection}
\caption{$Seg_1$ is the intersection between $S_2$ and $F_1$. $Seg_2$ is the intersection between $S_1$ and $F_2$. The intersection between $F_1$ and $F_2$, which is the line segment in red, is the overlap of $Seg_1$ and $Seg_2$.}
\label{fig_projection}
\end{figure}

M\"{o}ller's algorithm computes the intersection between two triangles $F_1$ and $F_2$ in three steps, as illustrated in Fig. \ref{fig_projection}. First, an early rejection is performed by testing whether $F_1$ intersects the plane where $F_2$ lies, which is denoted as $S_2$. This is a necessary condition of intersection between $F_1$ and $F_2$. If this test is passed, then the same is performed for $F_2$ and $S_1$, where $S_1$ is the plane of $F_1$. Second, the intersection between $F_1$ and $S_2$, denoted as $Seg_1$, and the intersection between $F_2$ and $S_1$, denoted as $Seg_2$, are separately computed . At last, the intersection between $F_1$ and $F_2$ is determined by computing the overlap area of $Seg_1$ and $Seg_2$ .

\begin{figure}[t]
  \centering
  % Requires \usepackage{graphicx}
  \includegraphics[width=3.0in]{nonrobust}\\
  \caption{The left edges of A and B (\emph{left}) are nearly but not exactly collinear. However, under usual float point arithmetic, they might be judged as collinear causing discontinuous edges in the final result (\emph{right}).}\label{fig:precision}
\end{figure}

Conventional implementations of M\"{o}ller's algorithm use vertex-based representation and usual float point arithmetic. However, this is neither exact nor robust. In Fig. \ref{fig:precision}, we illustrate a non-robust 2D case of the Boolean operation. Although implementing M\"{o}ller's algorithm with arbitrary precision arithmetic could make it robust, it is too costly for a large CSG evaluation. Actually, the non-robustness of this algorithm is from \emph{constructions} \cite{bernstein2009fast}, which compute new coordinates of geometry objects based on the known coordinates of existing ones. In M\"{o}ller's algorithm, the coordinates of $Seg_1$ and $Seg_2$ (Fig. \ref{fig_projection}) are computed as intermediate results. On the other hand, if the computation can be restricted to \emph{predicates}, which make a two or three-way decision based on known coordinates, we could quickly and robustly implement this algorithm.

Fortunately, according to Sugihara and Iri \cite{sugihara1990solid}, geometry computation of Boolean operations can be restricted to predicates using plane-based representation of polyhedra. This motivated us to integrate plane-based representation into M\"{o}ller's algorithm. Exact and efficient predicate computation is performed using precise geometry predicate techniques \cite{shewchuk1997adaptive}. Details of our implementation are shown below.

\begin{figure}
  \centering
  % Requires \usepackage{graphicx}
  \includegraphics[width=1.7in]{p-reps}\\
  \caption{Plane-based representation of triangles. The yellow triangle is represented by three bounding planes (green) and the supporting plane where the triangle lies.}\label{fig:p-reps}
\end{figure}

We first convert $F_1$ and $F_2$ into their respective plane-based representations: a supporting plane surrounded by three bounding planes (Fig. \ref{fig:p-reps}). We implement this conversion using the exact conversion method by Campen and Kobbelt \cite{campen2010exact}. Then, in our implementation, the relative position of a point and a plane can be determined by computing the sign of the multiplication of determinants \cite{sugihara1990solid}. To avoid introducing errors, we represent the end points of $Seg_1$ and $Seg_2$  by intersections of plane triples, rather than by directly computing their coordinates. As illustrated in Fig. \ref{fig:isect}(a), these points are the intersection between an edge line of a face and the supporting plane of the other face. Because the edge line of $F_1$ can be represented as $[S_1, S_X]$ (or $[S_2, S_X]$ for an edge line of $F_2$), where $S_X$ is the corresponding bounding plane of the edge, the end points of $Seg_1$ or $Seg_2$ can be represented in the form of $[S_1,S_2,S_X]$. In Fig. \ref{fig:isect}, we list all intersection situations between a triangle and a plane, and the $S_X$  in each situation.

\begin{figure}[t]
\centering
\includegraphics[width=3.5in]{sign}
\caption{We denote the signed distance from point $X$ to plane $S_2$ as $d_X$. All the four conditions of intersection between $F_1$ and $S_2$ (denoted as $Seg_1$) are:  (a) $d_A\cdot d_C<0$, $d_B\cdot d_C<0$; (b) $d_A=0$, $d_B=0$, $d_C\neq 0$; (c) $d_A=0$, $d_B\cdot d_C<0$; (d) $d_A=0$, $d_B\cdot d_C>0$. End points of $Seg_1$ are intersections between $S_2$ and related edge lines of $F_1$ (bold edges).}
\label{fig:isect}
\end{figure}

>>>>

\usepackage{enumitem}
%\usepackage[]{algorithm2e}
\usepackage[]{algpseudocode}
\usepackage{algorithm}
\usepackage{multirow}
\usepackage{slashbox}
\usepackage{color}
%\usepackage[numbers,sort&compress]{natbib}
%\usepackage{ulem}
%\usepackage[numbers,sort&compress]{natbib}
\renewcommand{\arraystretch}{1.3}

\usepackage{color}
\usepackage{bm}
%\usepackage{mathrsfs}
\usepackage[mathcal]{eucal}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{wrapfig}

\usepackage{amssymb}% http://ctan.org/pkg/amssymb
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%
% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\newcommand{\tabincell}[2]{\begin{tabular}{@{}#1@{}}#2\end{tabular}}

\floatname{algorithm}{Algorithm}

\begin{document}
\title{Efficient and Robust Plane-Based Boolean Operations on N Meshes}
\author{%Rui~Wang,~
        %Xudong~Jiang,~
        %Hongbo~Fu,~
        %Ping~Li,~
        %Bin~Sheng~and
        %Enhua~Wu
\IEEEcompsocitemizethanks{

@@

\IEEEtitleabstractindextext{
\begin{abstract}
  Constructive solid geometry (CSG) is widely used for computer aided design and manufacturing. However, boolean operations, which compute the boundary representation of CSG, have been suffered from robustness and efficiency problem for more than three decades. We propose a fast and exact boolean operation method which is unconditionally robust for regular set solids. We notice that among previous works,boolean methods using plane representations (P-reps) of meshes are easy to achieve robustness, but hard to be efficient. On the other hand, methods based on boundary representation (B-reps) of meshes are fast but not robust unless exact arithmetic is applied. Therefore, we choose to combine P-reps with B-reps, which allows us to take the advantages of both. Our method is globally based on B-reps for its efficiency, and P-rep geometry is applied locally to ensure robustness. Comparison experiments with state-of-the-art techniques show that our method is unconditionally exact and robust, while is competitively efficient as non-robust methods.
\end{abstract}

\begin{IEEEkeywords}
boolean operations, CSG evaluation, plane-based geometry
\end{IEEEkeywords}

@@

\IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
\IEEEPARstart{C}{onstructive} solid geometry (CSG) has long been a popular modeling tool of computer-aided design and computer-aided manufacturing (CAD/CAM). It constructs complex models by combining primitives using a series of regularized boolean operations \cite{requicha1977mathematical,tilove1980closure}: union, intersection and difference. CSG can be converted into the widely-used boundary representation (B-rep) through boolean operations, which is an classical topic with history of over three decades.

In general, there are two major types of boolean algorithms according to how they deal with intersection between primitive meshes. Approximate methods \cite{wang2011approximate,pavic2010hybrid,biermann2001approximate} redicretize the intersection areas, fit vertices approximatively and rearrange topology. On the other hand, exact method do not change vertex positions, retain as many input elements  (e.g., faces, vertices and topology) as possible. In many applications, exact methods are preferred for their accuracy. Also, using exact methods, the clear mapping between the surface of input and output meshes makes it easier for the inheritance of information like face colors and materials.

However, for exact boolean operation methods, there is always a bargain between robustness and efficiency. Many exact boolean operation methods focus on efficiency. But for robustness, the solutions are either exact arithmetic \cite{barki2015exact,zhou2016mesh} or trivial workarounds to avoid degenerations, such as setting a tolerance \cite{feito2013fast,segal1990using} and rotating the primitive with small angles \cite{douze2015quickcsg}. Methods using exact arithmetic are significantly slower, and workarounds are often case dependent and not reliable. Recently, some methods \cite{bernstein2009fast,campen2010exact} seek unconditional robustness by avoiding geometry constructions (such as computing new vertex coordinates) and using only predicates. Their key idea is to utilize plane-based representations (P-reps) of meshes instead of boundary representation (B-reps), which comes from the observation of Sugihara and Iri \cite{sugihara1990solid} that boolean operations can be performed without geometry constructions if P-reps are used. While these plane-based methods are generally faster than exact arithmetic ones, they still suffer from performance issue because of the complex conversions among different representations. Also, the incoherence between different representations requires extra steps to reconstruct geometry connectivity, leading to extra computation burden and more complex topology.

Inspired by these previous work, we develop an unconditionally robust boolean operation method which is as robust as P-rep-based method and as fast as non-robust B-rep-based method. To achieve this, we embed P-rep information into faces of B-reps, forming hybrid representation of meshes. We avoid geometry constructions throughout our method, and convert all necessary computation into predicates under the hybrid representation. In general, the B-reps information is used for coarse tests and efficient face neighbor query, while P-reps information is for exact geometry computation. During intersection computation, we encode the intersections between meshes into sets of planes and use these planes to perform exact tessellation. After that, we classify each faces in the tessellated meshes using local binary space partition (BSP) trees, which also facilitate the previous plane-based representation to guarantee exactness.  In addition, while most previous boolean operations can only deal with two meshes, our method can perform boolean operations with arbitrary meshes in one call, instead of recursively doing incremental boolean operations. In this way, our method could save computation time by avoiding repetitive computation. The experiments show that our method is much faster than existing robust exact methods, while only about two times slower than the state-of-art non-robust exact methods.

@@

Boolean operations have been researched for over three decades since it inception in the 1980s \cite{requicha1985boolean, laidlaw1986constructive}. Previous works can be classified into two categories: exact methods and approximate methods. Exact methods retain vertex positions. Also the topology of the input meshes is preserved as much as possible. The coordinates of intersection points between input meshes are computed by input configurations to construct the topology around intersection. However, the coordinates cannot be represented exactly using fix-precision float-point numbers and the robustness is always a problem. On the other hand, approximate methods rediscretize the input mesh surfaces using techniques such as voxels, octrees and Layered Depth Images. They can reach better performance and robustness, but the loss of geometry information and precision is inevitable.

\subsection{Exact Methods}

\iffalse
\begin{figure}[t]
\centering
\includegraphics[width=2.5in]{coplanarexample}
\caption{{\color{red}{Sketch: the example of coplanar cases in a boolean operations with four cylinders}}}
\label{fig:coplanarexample}
\end{figure}
\fi

For exact methods, some \cite{ogayar2015deferred,douze2015quickcsg,zhou2016mesh,xu2013fast,feito2013fast} pursues efficiency but cannot guarantee robustness in degenerate cases. Most of them are implemented by fix-precision float-point arithmetic and numerical errors are inevitable. Douze et al. \cite{douze2015quickcsg} implemented very efficient method that can handle very large meshes. Also, they can perform non-incremental boolean operations: evaluating the final result mesh of CSG all at once instead of decomposing the CSG into a series of incremental boolean operations. However, their method can neither deal with coplanar cases nor guarantee accuracy of topology. In fact, large CSGs are more complex and more likely to contain degenerate cases, and more vulnerable under numerical errors. Efficiency only is far from being practical.

Many researchers have attempted to solve robustness issue for exact boolean operations using arbitrary precision arithmetic \cite{banerjee1996topologically, fortune1995polyhedral, keyser2004esolid, granados2003boolean, hachenberger2005boolean} and exact interval computation \cite{fang1993robustness, hu1996robust, segal1990using}. However, these methods are too expensive in computation time and memory to be practical for large CSGs. For example, one of the available implementation of robust exact boolean operations method is CGAL's \cite{cgal:hk-bonp3-15a} exact-arithmetic implementation \cite{granados2003boolean} of Nef polyhedra \cite{bieri1988elementary}. It requires conversions between Nef polyhedra and input mesh data structures during evaluation and is more than 50 times slower than non-robust booleans \cite{bernstein2009fast}.

\subsection{Approximative Methods}

Since efficiency, accuracy and robustness are hard to be satisfied at the same time, some method choose to sacrifice accuracy for better efficiency and robustness. Most of such methods are based on conversion to volumetric representations. However, the quality of result mesh depends on the resolution of the volume grid and better quality requires dramatically higher resolution. To accelerate this process, some try to reduce the complexity of the output mesh \cite{varadhan2004topology} and others try to preserve non-intersected areas of the input mesh to avoid redundant tessellation \cite{pavic2010hybrid,wang2011approximate,zhao2011parallel,hable2005blister,ogayar2006gpu}. Also, with the development of general-purpose computing on graphics processing units (GPGPU), many of them utilize the grand computation power of graphics hardware for boolean operations. These methods have good performance and are suitable for interactive applications. However, the fundamental problem of approximate methods is owing to grid-depend nature of volumetric calculations: they inevitably suffers from geometric detail loss and unwieldy topology changes.

\subsection{Plane-Based Methods}
\label{sec:pbrelated}

The concept of plane-based representations (P-reps) of polygonal meshes was first described by Sugihara and Iri \cite{sugihara1990solid}. P-reps provide essential benefit for boolean operations that no new geometry primitive has to be constructed to obtain the results---they can be composed of a subset of the planes from the input polyhedra. It means the computation can include only geometry predicates, which are much easier to be implemented exactly and robustly. Based on P-reps, Berstein and Fussell \cite{bernstein2009fast} combined the P-reps with binary space partition (BSP) trees \cite{thibault1987set,naylor1990merging} to develop a boolean operation method which is unconditionally robust under consistent inputs . Compared to B-rep-based methods, their method does not rely on exact arithmetics to be robust. However, the merge of two BSP trees is an $O(mn)$ time complexity process, where $m$ and $n$ are the size of the trees. It makes this method impractical for large meshes. Later, Campen and Kobbelt \cite{campen2010exact} improved their method by localize BSP operations using an octree. The mesh refinement only takes place locally near intersections, which is much faster than Berstein and Fussell's work.

\section{Background and Overview}

\label{sec:overview}
Our method is designed for boolean operations on arbitrary number of regular triangle meshes \cite{requicha1985boolean}. Input meshes are required to be free from self-intersecting, and they are represented with B-reps with geometry connectivity (e.g., halfedge, winged-edge, etc.). Despite of these requirements, our method is robust and not sensitive to topology deficiencies like holes and self-intersections and works correctly if deficiencies are not near the intersection areas between meshes.

\subsection{Boolean Numerical Errors}
\label{sec:paradigm}

The boolean operation of triangle meshes is in essence a process of face selection. Namely, given a boolean function, it reserves those primitive faces that pass the function and drop the others to generate the final mesh. Unfortunately, for B-rep meshes, not all the input faces can be classified as a whole, since for faces near the intersections, only parts of them belong to the final mesh. Therefore, we need an extra step aiming at detecting intersections between meshes and performing tessellation. Most of the boolean operation methods follow this two-step paradigm, which consists of intersection computation and face classification. So is our method.

The first step is intersection computation. Input faces are tested in pairs to compute their intersections. Each input meshes are tessellated according to the intersection results, making every face be completely inside, outside or on the boundary of other primitives. The tessellated meshes are what we called \emph{intersection-free meshes}. Unfortunately, under fix-precision float-point arithmetic, intersection tests are error-prone: first, degenerate cases of intersection are hard to detect; second, when there are intersections, new vertices are introduced into the geometry, whose coordinates cannot be exactly represented by fix-precision float-point numbers.

The second step is face classification. We evaluate whether a given face $\bm{s}$ belongs to the final geometry according to the $n$-primitive boolean function $f$ :
\begin{equation}
\lambda_f(\bm{s}) = f(\boldsymbol{\Lambda}(\bm{s})) = f(\lambda_1(\bm{s}), \lambda_2(\bm{s}), \cdots, \lambda_n(\bm{s})).
\end{equation}
The parameter $\lambda_i(\bm{s})$ is space indicator with respect to mesh $M_i$. To compute $\lambda_f(\bm{s})$ and classify $\bm{s}$, one has to know the space indicators with respect to all the primitives ($\boldsymbol{\Lambda}(\bm{s})$). A space indicator has four conditions: completely inside (\emph{in}), completely outside (\emph{out}), on the boundary with consistent normals (\emph{same}) or opposite normals (\emph{oppo}). The rules of boolean function evaluation are described in \cite{douze2015quickcsg,feito2013fast}.

\begin{figure}[t]
\centering
\includegraphics[width=2.2in]{boolean-01}
\caption{In the 2D view, face $\bm{s}$ (the red line segment) from mesh $M_2$ is on the surface of mesh $M_1$. However, because we use face barycenter $\bm{v}_{c(\bm{s})}$ to compute the indicator, whose coordinates contains round-off error, the point could  move to $\bm{v'}_{c(\bm{s})}$. Then we might falsely classify $\bm{s}$ as outside of $M_1$.}
\label{fig:falseclass}
\end{figure}

Face classification also suffers from numerical errors. Many methods classify face according to the indicators of its barycenter using point-in-polyhedron test \cite{feito2013fast,campen2010exact}. However, coordinates of barycenters cannot be exactly represented and could generate false classification (illustrated in Fig. \ref{fig:falseclass}). What makes the problem worse is that for the large amount of faces and input meshes, many methods \cite{pavic2010hybrid,feito2013fast,ogayar2015deferred,zhou2016mesh} take the benefits of the local coherence of indicators, classifying neighboring faces together . Despite of the performance improvement, it can propagate false classification to neighboring faces and lead to wide-range failure.

\subsection{Plane-based Geometry}

\begin{wrapfigure}{r}[0in]{0in}
\includegraphics[width=1.5 in]{p-reps}
%\caption{Plane-based representation of triangle}
\end{wrapfigure}

In our method, we avoid numerical errors by embeding P-reps into conventional B-rep-based boolean algorithms and substituting constructions under B-reps with predicates under P-reps. Using P-reps, each face $\bm{s}$ with $n$ edges is represented by a supporting plane $\bm{p}_{s,sp}$, where the face lies, and a list bounding planes $\{\bm{p}_{s,b}^i \ \vert\  i = 0, 1,...,n-1\}$. Each edge line $\bm{e}_{\bm{s}}^i$ is represented by intersection $\bm{p}_{s,sp} \cap \bm{p}_{s,b}^i$. Corner vertex $\bm{v}_s^i$ is represented by $\bm{p}_{s,sp} \cap \bm{p}_{s,b}^i \cap \bm{p}_{s,b}^{{(i+1)}\bmod{n}}$. We use Campen et al.'s method \cite{campen2010exact} for exact conversion of triangles to their P-reps. All geometry predicates are computed using filtering techniques proposed by Shewchuk \cite{shewchuk1997adaptive} for both exactness and efficiency.

Other commonly used notations in this paper are presented here.
%A plane $\bm{p}$ can be represented by four scalar coefficients $a(\bm{p}), b(\bm{p}), c(\bm{p}), d(\bm{p})$.
The normal of a plane $\bm{p}$ is denoted as $\bm{n}(\bm{p})$. A line $\bm{l}$ can be represented by intersection of two planes $(\bm{p}_l^0 \cap \bm{p}_l^1)$, or in short, $\bm{l}\colon(\bm{p}_l^0 \cap \bm{p}_l^1)$. The positive direction of the line $\bm{l}$ is defined by $\bm{n}(\bm{p}_l^0) \times \bm{n}(\bm{p}_l^1)$. A point $\bm{v}$ can be represented by non-trivial plane triples $(\bm{p}_v^0 \cap \bm{p}_v^1 \cap \bm{p}_v^2)$, or in short, $\bm{v}\colon(\bm{p}_v^0 \cap \bm{p}_v^1 \cap \bm{p}_v^2)$.

\iffalse
Given that the input vertex coordinate has $L$-bit precision including sign, the precisions of the four plane coefficients are $L_a, L_b, L_c, L_d$, which have to satisfy the following requirements:
\begin{equation}
\begin{split}
&L_a=L_b=L_c\ge 2(K-1)+1+1,\\
&L_d\ge(L_a-1)+(L-1)+2+1.
\end{split}
\end{equation}
Here $K$ is the maximum bits required to represent edge vector components in input meshes. If we use $M$ bits to store each plane coefficient, basically $M$ has to be greater than $L_d$. In our implementation, we actually use a more strong constraint $M \ge L_d+1$. Under this constraint, given a $L$-bit vertex $\bm{v_i}$, we can compute the orientation of plane $p_j$ with respect to $\bm{v_i}$ by the sign $\bm{n}(\bm{p_j})\cdot\bm{v_i} + d(\bm{p_j})$, which requires no more than $M$ bits. It allows very fast plane orientation predicates which accelerates the early rejection in our triangle-triangle intersection tests (\S \ref{section:isect}).

Let $\delta = 2^{K-L-1}$ be the relative length (in max-norm) of the longest edge in the mesh (relative to the bounding box). If we use IEEE 754 double float point number ($M=53$) and assume $\delta \le 2^{-5.5} \approx 0.022$, the maximum input precision $L$ can reach to 20, which is enough for most application (for standard IEEE 754 single float point number, the precision is 24-bit).
\fi

\label{sec:substrates}
Plane-based geometry predicates used in our method are mostly well-discussed in previous work \cite{bernstein2009fast,banerjee1996topologically}. In the following, we focus on two predicates related to sorting, which are designed for our method particularly.

\vspace{0.5em}
\noindent \textbf{Linear ordering of points}

\begin{figure}
  \centering
  % Requires \usepackage{graphicx}
  \includegraphics[width=3.5in]{twopointoneline}\\
  \caption{Geometry configuration of linear ordering of points. Points $\bm{v}_a$ and $\bm{v}_b$ are both on line $\bm{l}_{ab}$. We convert this problem into the plane ordering of $\bm{p}_a$ and $\bm{p}_b$ along $\bm{l}_{ab}$.}\label{fig:twopointoneline}
\end{figure}

\noindent Given an line $\bm{l}_{ab}$ with two points $\bm{v}_a\colon(\bm{p}_a^0\cap\bm{p}_a^1\cap\bm{p}_a^2)$ and $\bm{v}_b\colon(\bm{p}_b^0\cap\bm{p}_b^1\cap\bm{p}_b^2)$ on it, we need to determine the linear order of the two points along $\bm{l}_{ab}$ (Fig. \ref{fig:twopointoneline}). To solve this problem, we choose one plane not parallel with $\bm{l}_{ab}$ from the plane representation of each point and convert this problem into linear order of planes, which can be solved by Banerjee et al.'s method \cite{banerjee1996topologically}. The chosen planes should have the same orientation with $\bm{l}_{ab}$ (the dot product between plane normal and $\bm{l}_{ab}$ has to be positive) and unqualified planes have to be flipped.

\begin{wrapfigure}{r}[0in]{0in}
\includegraphics[width=1.5 in]{boolean-02}
%\caption{Plane-based representation of triangle}
\end{wrapfigure}
\vspace{0.5em}
\noindent \textbf{Circular ordering of lines}~~~~

\noindent During face tessellation, we need to know the which edges are neighboring (Fig. \ref{fig:circularorder}), that is, circular ordering of directed lines around a vertex. Lines can be sorted in a divide-and-conquer way based on the relative order of each pair of lines. Therefore, this problem is converted to that given two lines $\bm{l}_a$ and $\bm{l}_b$ in a plane $\bm{p}_0$ and their intersection point $\bm{v}_{ab}$, compute the circular order of the two lines. We can compute the order by the sign of $\sin{\theta_{ab}}$, where $\theta_{ab}\in(-\pi,\pi)$ is the angle from $\bm{l}_a$ to $\bm{l}_b$ under the top view of $\bm{p}_0$. The sign of $\sin{\theta_{ab}}$ is the same as triple product $\bm{n}(\bm{p}_0) \cdot (\bm{l}_a\times\bm{l}_b)$. However, explicit computing this equation is complicated because both $\bm{l}_a$ and $\bm{l}_b$ are implicitly represented by intersection of planes.
